---
title: "REPORT/PROJECT TEMPLATE"
author: "Julia Cancro"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y %H:%M')`"
output:
  html_document:
    toc: true
    toc_float: true
    keep_md: true
    self_contained: true
editor_options: 
  chunk_output_type: console
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
---

THIS TEMPLATE IS DESIGNED TO HELP GUIDE YOUR PROJECT DEVELOPMENT. YOU CAN CHANGE
OR ADD ANY SECTIONS, BUT I NEED A BRIEF BACKGROUND (WHAT IS YOUR QUESTION, WHY,
WHAT DOES THE DATA LOOK LIKE?). THEN YOU NEED TO IMPORT THE DATA, CHECK IT, PERFORM
EDA, RUN AND INTERPRET TESTS, DEVELOP APPROPRIATE PLOTS, AND THEN BRIEFLY SUMMARIZE
YOUR FINDINGS.

# Background and metadata

What is the main goal of your project? How did you collect the data? You can copy
this to/from a paper.

- The goal of this experiment is to quantify the impact of microzooplankton grazing on phytoplankton assemblages within NYC embayments.

# Read in (Import) your data

Useful functions here may be **read.csv**.

```{r}
JBWR1 <- read.csv("~/Downloads/JBWR1.csv", stringsAsFactors = T)
#need to use stringsAsFactors to make characters read in as factors
```

# Check your data
```{r}
head("~/Downloads/JBWR1.csv")
```
Useful functions here may be **str**, **summary**, or **head**.

# Exploratory data analysis

Install and call the package

```{r}
library(ggplot2)
```

To make a plot, set a base layer using the **ggplot** fuxn
Have to add geom below because otherwise the fxn is not defined

```{r}
P.Growth_Dilution_Relationship <- ggplot(JBWR1, aes(x=Dilution_Factor, y = Net_Growth_Rate))
```

Here we are naming a dataframe to use (first argument), then we note which 
columns to use for the x and y axis

When we do this, we get a blank graph, but we have to name the ggplot to see it

```{r}
P.Growth_Dilution_Relationship
```

We can create a scatter plot using the geom_point command.
```{r}
P.Growth_Dilution_Relationship_scatter <- ggplot(JBWR1, aes(x=Dilution_Factor, y = Net_Growth_Rate)) + 
  geom_point()
```

```{r}
P.Growth_Dilution_Relationship_scatter
```

```{r}
ggplot(JBWR1, aes(x=Dilution_Factor, y = Net_Growth_Rate)) +
  geom_point()
```

Useful functions here may come from the **ggplot2** package.

# Run statistical tests

We will typically focus on linear models. The most useful function here is
**lm**; it covers most tests from classical statistics.

The assumptions for all linear models 
are that the residuals are independently and identically distributed and follow 
a normal distribution.

Once you create the **lm** object, check it for assumptions using **plot(object)**.
What you want to see 

* is no pattern in first graph (Residuals vs fitted)
  * a pattern (like a funnel shape) means the residuals are increasing with the mean value, so you may need a transformation
* the dots falling on the line in the second graph (the Normal Q-Q).  this is a qq plot; if the lines fall along the line it means the normality assumption is ok
* no outliers in the 4th plot (Residuals vs Leverage); this identifies plots that are high leverage (removing/including them really changes your results). you may want to check these.

Despite all this, lm's are very robust and are often used in ecology.

## ANOVA example (comparing numerical data from categories)

```{r}
#example of anova
sleep <- read.csv("https://raw.githubusercontent.com/jsgosnell/CUNY-BioStats/master/datasets/sleep.csv", stringsAsFactors = T)
sleep$Exposure <- factor(sleep$Exposure)
sleep_by_exposure_lm <- lm(Dreaming~Exposure, sleep)
plot(sleep_by_exposure_lm)
library(car)
Anova(sleep_by_exposure_lm, type = "III")
```
If you get a significant p-value and you have more than 2 levels, an obvious 
question is which levels are different. Talk to me about exploring this with

```{r}
library(multcomp)
comparison <- glht(sleep_by_exposure_lm, linfct = mcp(Exposure = "Tukey"))
summary(comparison)
```

## Regression example (looking for relationships among numerical data)


```{r}
#example of regression
dreaming_by_sleep_lm <- lm(Dreaming~TotalSleep, sleep)
plot(dreaming_by_sleep_lm)
Anova(dreaming_by_sleep_lm, type = "III")
#summary is good to get the coefficients
summary(dreaming_by_sleep_lm)
```

# Publication-quality graphs

# Brief discussion